Deciding Bitvector Termination with Program Synthesis

### 2014-07-02 papers/termination/kalashnikov.tex
modified kalashnikov + other small things
revising sec 6

\subsection{The Basic Synthesis Algorithm}
\label{sec:abstract-algorithm}

@@

When the quantifier free first-order fragment of the logic is efficiently decidable and
$P$ can be expressed as a ground term of the logic, we can use
Counterexample Guided Inductive Synthesis (CEGIS)~\cite{lezama-thesis,sketch} to
check the validity of the synthesis formula.  The core of the CEGIS algorithm is
a refinement loop to check the
validity of the synthesis formula as shown in Fig.~\ref{fig:abstract-refinement} and
Algorithm~\ref{fig:abstract-refinement-code}.  The algorithm is divided into two
components: {\sc synth} and {\sc verif}.  We track a finite (small) subset of test inputs
and {\sc synth} synthesises a candidate program that is correct on just those tests.
{\sc Verif} then tries to verify the candidate program by searching for an input on
which the candidate program does not satisfy the specification.  If no such input
can be found, the candidate program is correct.  Otherwise, the new input is added
to the set we must check in {\sc synth}.  If the input set is finite, this
procedure is guaranteed to terminate.

@@

\subsection{The Concrete Algorithm for Bitvector Programs}
\label{sec:concrete-algorithm}

@@

To synthesise tricky bitvector programs like this, it is natural for us to
work in the logic of quantifier-free propositional formulae and to use a
propositional SAT or SMT-$\mathcal{BV}$ solver as the decision procedure. 
However we propose a slightly different tack, which is to use a decidable
fragment of C as a ``high level'' logic.

\subsubsection{\newC}
\label{sec:logic}

@@

\subsubsection{Encoding an $\mathcal{L}$-Program in \newC}
\label{sec:encode-l}

>>>>

\iffalse
\subsection{The Basic Synthesis Algorithm}
\label{sec:abstract-algorithm}

@@

In this section we present a solver for the synthesis fragment. An
important characteristic of our solver is that is models bit accurate
semantics. This enables the precise treatment of bit-wise operations and
arithmetic modulo fixed widths.

%We can solve the synthesis problem using any off-the-shelf program synthesiser.  If you don't
%have a program synthesiser kicking around, we present the design of one synthesiser
%that can be used to solve the problem.
%We start by recalling the synthesis formula:
% $$\exists P, x . \forall y . \sigma(P, x, y)$$
We start by recalling the synthesis fragment:
 \[
  \exists P,~ x . ~\forall~ y . \sigma(P, x, y)
 \]

where $P$ ranges over sets, while $x$ and $y$ range over ground terms (we have dropped
 the vector symbols for brevity reasons).

When the quantifier free first-order fragment of the logic is efficiently decidable and
$P$ is expressible as a ground term of the logic, we can use
Counterexample Guided Inductive Synthesis (CEGIS)~\cite{lezama-thesis,sketch} to
check the validity of the synthesis formula. The core of the CEGIS algorithm is
the refinement loop in Fig.~\ref{fig:abstract-refinement} and
detailed in Algorithm~\ref{fig:abstract-refinement-code}.

The algorithm is divided into two
components: {\sc synth} and {\sc verif}. We track a finite set of test
inputs and {\sc synth} synthesises a candidate program that is correct on just those tests.
{\sc Verif} then tries to verify the candidate program by searching for an input on
which the candidate program does not satisfy the specification.  If no such input
can be found, the candidate program is correct.  Otherwise, the new input is added
to the set we must check in {\sc synth}.  If the input set is finite, this
procedure is guaranteed to terminate.

@@

\iffalse
\subsection{The Concrete Algorithm for Bitvector Programs}
\label{sec:concrete-algorithm}

@@

To synthesise tricky bitvector programs like this, it is natural for us to
work in the logic of quantifier-free propositional formulae and to use a
propositional SAT or SMT-$\mathcal{BV}$ solver as the decision procedure. 
However we propose a slightly different tack, which is to use a decidable
fragment of C as a ``high level'' logic.
\fi

\subsection{\newC}
\label{sec:logic}

@@

%\subsubsection{Encoding an $\mathcal{L}$-Program in \newC}
\label{sec:encode-l}
